Numerical modelling of the Cordilleran ice sheet

### 2014-07-19 kappa.tex
Add real conclusions.

Certainly the most important assumption made in this thesis is that of a
Cordilleran ice sheet largely disconnected from its much larger neighbour, the
Laurentide ice sheet. Geomorphological evidence from the suture zone indicates
that, during the LGM, both Cordilleran and Laurentide ice flow has been
deflected by their junction \citep{Kleman.etal.2010, Margold.etal.2013,
Margold.etal.2013a}, although it remains unclear for how long this junction has
lasted \citep{Jackson.etal.1997, Bednarski.Smith.2007, Gowan.2013}.

@@

Another important source of error is the treatment of precipitation, to which
the numerical ice sheet model is highly sensitive (\CCLI). In all sensitivity
studies, we base precipitation patterns on current values derived atmospheric
reanalyses. Although the distribution of precipitation has undoubtedly
been affected by past changes in global climate and the intermittent presence
of ice sheets and ice caps in the Northern Hemisphere, its changes are
difficult to assess. In the North American Cordillera, much of the continental
interior was likely dryer than today due to the presence of ice caps on the
Pacific Coast ranges, reinforcing the current topographic barrier. However,
some regions were probably rainier as a side effect of modified wind patterns.

@@

Geothermal heat flux forcing, a major contributor to the modelled distribution
of en-glacial temperatures, and the location of basal melt areas, was assumed
constant over the modelling domain. However, the geology of the study area
consists of an amalgamation of former subduction zones, resulting in complex
geothermal patterns \citep{Blackwell.Richards.2004}. I am now at the stage to
start including this variability as a driver to the ice sheet model, which will
offer new opportunities to study its output in terms of ice flow patterns,
consistent basal freezing and basal melting patterns, regions prone for glacial
erosion, and regions characterised by glacial deposition.

@@

>>>>

Certainly the most important assumption made in this thesis is that of a
Cordilleran ice sheet largely disconnected from its much larger neighbour, the
\emph{Laurentide ice sheet}. Geomorphological evidence from the suture zone
indicates that, during the LGM, both Cordilleran and Laurentide ice flow has
been deflected by their junction \citep{Kleman.etal.2010, Margold.etal.2013,
Margold.etal.2013a}, although it remains unclear for how long this junction has
lasted \citep{Jackson.etal.1997, Bednarski.Smith.2007, Gowan.2013}.

@@

Another important source of error is the treatment of \emph{precipitation}, to
which
the numerical ice sheet model is highly sensitive (\CCLI). In all sensitivity
studies, we base precipitation patterns on current values derived atmospheric
reanalyses. Although the distribution of precipitation has undoubtedly
been affected by past changes in global climate and the intermittent presence
of ice sheets and ice caps in the Northern Hemisphere, its changes are
difficult to assess. In the North American Cordillera, much of the continental
interior was likely dryer than today due to the presence of ice caps on the
Pacific Coast ranges, reinforcing the current topographic barrier. However,
some regions were probably rainier as a side effect of modified wind patterns.

@@

\emph{Geothermal heat flux} forcing, a major contributor to the modelled
distribution
of en-glacial temperatures, and the location of basal melt areas, was assumed
constant over the modelling domain. However, the geology of the study area
consists of an amalgamation of former subduction zones, resulting in complex
geothermal patterns \citep{Blackwell.Richards.2004}. I am now at the stage to
start including this variability as a driver to the ice sheet model, which will
offer new opportunities to study its output in terms of ice flow patterns,
consistent basal freezing and basal melting patterns, regions prone for glacial
erosion, and regions characterised by glacial deposition.

@@

Through sensitivity studies, and comparison of model output to geological
evidence, I have gained a personal understanding of Cordilleran ice sheet
evolution through the last glacial cycle. I attempt to summarise my thoughts
here in a few points, perhaps also applicable to other ice sheets in the
Northern Hemisphere.

\begin{enumerate}

\item Although spatial patterns of glaciation in the North American Cordillera
are sometimes paradoxical, they can be largely explained by the present-day
distribution of temperature and precipitation in the region, provided an
accurate treatment of seasonality, orographic precipitation, and daily
variability in the model, and a simple temperature-index accumulation and melt
model.

\item The LGM Cordilleran ice sheet corresponds to a transient stage, out of
balance with the contemporaneous atmospheric temperatures. Its maximum extent
is the result of a 50\,ka-long evolution, captured in the geologic record at a
moment of transition from colder to warmer temperatures than those needed to
sustain a constant ice sheet geometry. Therefore, this maximum stage is not
representative of average conditions of the last glacial cycle.

\item The patterns of growth and decay of the Cordilleran ice sheet are varied
and differ by region and by epoch. Two-dimensional, conceptual models of growth
and decay \citep[e.g.][]{Fulton.1991, Margold.etal.2013a} do not encompass this
diversity. In order to decipher the complexity of geological evidence, the
Cordilleran ice sheet needs to be thought of as a three-dimensional body of
moving ice in perpetual transformation towards evolving climate conditions.

\end{enumerate}
